Multi-Resolution Spatiotemporal Neural Field Estimation

### 2011-03-08 Multiresolution_framework/ltx/manuscript.tex
edit to discussion

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Multi-Resolution Spatio-Temporal Neural Field Estimation }

@@

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{equation}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(s)=[
    \boldsymbol\phi_{j_0}^\top(s) ~
    \boldsymbol\psi_{j_0}^\top(s) ~ 
    \boldsymbol\psi_{j_0+1}^\top(s) ~
    \cdots ~
    \boldsymbol\psi_{j}^\top(s)]
\end{equation} 
\begin{equation}
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (r) = [
    \boldsymbol\phi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0+1}^\top(r) ~ 
    \cdots ~ 
    \boldsymbol\psi_{j}^\top(r)
]   
\end{equation}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 200 realizations of 1~s of data were generated, where the estimation was applied to the final 900~ms, allowing the model's initial transients to die out. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The firing rate slop, $\varsigma$ was set to $0.56~mV^{-1}$. The distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed.  %\ken{if we need to gain space I would remove the rest of this paragraph and put all parameters in a table} 
The spatial cutoff frequency of the observed field, \parham{$y_t(r_n)$}, was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, \parham{this was obtained by averaging the spectral power of the spatial frequency (over time) from the observations} \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.

@@

  \mike{The framework makes two key assumptions about the cortex: a linear activation function and stationary dynamics. Our main aim for this framework in the future is to relax these assumptions using the multi-resolution decomposition. While the decomposition itself holds, efficiently performing nonlinear smoothing in the much larger state space remains a challenge. Additional avenues for future work include extending the model to capture a second order synaptic response kernel, and time delays \parham{at lower spatial resolutions}.  The estimation techniques should also be extended to deal with spatial and temporal heterogeneity in the kernel. Finally, the majority of our current work involves applying this framework to collected data, primarily for the use of seizure detection and mitigation.}

 In order to apply the framework to real data some assumptions must be made, such as stationarity of the parameters over the estimation period. Perhaps the most critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model \parham{and its ability to approximate the neural field at required resolution.} \cut{The decomposition holds for more sophisticated nonlinear forms of the field equations, where the unscented RTS smoother should be used in the E-step of the estimation algorithm.}

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Multi-Resolution Spatiotemporal Neural Field Estimation }

@@

The vectors of the kernel and the field scaling and wavelet functions, $\boldsymbol\lambda$ and $\boldsymbol\mu$, respectively, are defined by
\begin{align}
    \label{KernelBasisVector}
    \boldsymbol\lambda^\top(s)=&\left[
    \boldsymbol\phi_{j_0}^\top(s) ~
    \boldsymbol\psi_{j_0}^\top(s) ~ 
    \boldsymbol\psi_{j_0+1}^\top(s) ~
    \cdots ~
    \boldsymbol\psi_{j}^\top(s)\right] \\
    \label{FieldBasisVector}
    \boldsymbol\mu^\top (r) =& \left[
    \boldsymbol\phi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0}^\top(r) ~ 
    \boldsymbol\psi_{j_0+1}^\top(r) ~ 
    \cdots ~ 
    \boldsymbol\psi_{j}^\top(r)
\right]   
\end{align}

@@

\section{Results}
To demonstrate the performance of the MRAIDE estimation framework, data was generated synthetically  using \eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation}, enabling a comparison between true and estimated parameters. 200 realizations of 1~s of data were generated, where the estimation was applied to the final 900~ms, allowing the model's initial transients to die out. The observation noise was set to $\boldsymbol\Sigma_{\epsilon}=0.1 \times \mathbf{I}_{n_y}$ and the disturbance covariance was set to $\eta(r-r') = \phi_{3,-2}(r-r')$ (B-spline, $j=3$). The sampling period and the synaptic time constant were set to $T_s = 1$~ms and $\tau = 10$~ms, respectively. The firing rate slop, $\varsigma$ was set to $0.56~mV^{-1}$. The distance between adjacent sensors was $0.5$~mm, resulting into $n_y = 161$ observations. The observation kernel was modeled using a B-spline function, with a width of 0.8~mm at half the maximum amplitude. The spacing and bandwidth of the sensors allowed for the full spatial bandwidth of the field to be observed.  %\ken{if we need to gain space I would remove the rest of this paragraph and put all parameters in a table} 
The spatial cutoff frequency of the observed field was used to specify the level of decomposition required to represent the field using the wavelets. Using an oversampling parameter of 5, the cutoff spatial frequency was $\nu_{cy} \approx 6.9 $ cycles/mm, \parham{this was obtained by averaging the spectral power of the spatial frequency (over time) from the observations} \cite{Scerri2009}. Therefore, wavelets up to level $j=3$ (with the bandwidth $\approx[5,8]$ cycles/mm) can represent the significant spatial characteristics of the field, yielding $n_x = 131$ states.

@@

  % \mike{The framework makes two key assumptions about the cortex: a linear activation function and stationary dynamics. Our main aim for this framework in the future is to relax these assumptions using the multi-resolution decomposition. While the decomposition itself holds, efficiently performing nonlinear smoothing in the much larger state space remains a challenge. Additional avenues for future work include extending the model to capture a second order synaptic response kernel, and time delays \parham{at lower spatial resolutions}.  The estimation techniques should also be extended to deal with spatial and temporal heterogeneity in the kernel. Finally, the majority of our current work involves applying this framework to collected data, primarily for the use of seizure detection and mitigation.}

In order to apply the framework to real data some assumptions must be made, such as stationarity of the parameters over the estimation period. A critical assumption is that the model provides an apt description of the cortical dynamics. The authors acknowledge that there is, and will always be, a discrepancy between the model and cortex. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The key development in this paper is the multi-resolution decomposition forming the state-space model \cut{and its ability to approximate the neural field at required the level of accuracy.} While the decomposition holds for more sophisticated models, efficiently performing nonlinear smoothing in the much larger state-space remains a challenge. Additional avenues for future work include extending the model to capture a $2^{nd}$-order synaptic response kernel, and time delays according to spatial resolutions. The estimation techniques should also be extended to deal with spatiotemporal heterogeneity in the kernel. Finally, future work should be directed towards applying and validating the framework on real data.
